---
title: "Scatterplot Matrix"
author: "Dr. Weber"
date: "October 4, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(GGally)
library(Sleuth3)
```

## GGally Package

We will need an extension of the ggplot2 package called GGally. Now, we can use `ggpairs()` to create our scatterplot matrix. The advantage of this package is that it allows us to compare both numerical and categorical variables.

```{r}
mammals <- case0902
ggpairs(data = mammals, columns = 2:5)
```

Now, let's transform the data and look at the plots.

```{r}
mammals_trans <- log(mammals[,2:5])
ggpairs(data = mammals_trans)
ggplot2::ggsave("Notes_9_scattermatrix.png",path = "Figures")
```

Transforming the data allows us to see a pronounced relationship between each of our variables and our response. It also shows us that our explanatory variables are related to each other. This suggests we should fit a regression model for the log of brain weight on the log of each of these explanatory variables. 

```{r}
fit <- lm(Brain ~ Body + Gestation + Litter, data = mammals_trans)
summary(fit)
```