---
title: "Stat 361 - Simple Linear Regression"
author: "Dr. Weber"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(UsingR)
library(tidyverse)
```


## Simple Linear Regression

As a side note, this is old, clunky code: tough to read and understand what it is doing.

```{r}
freqData <- as.data.frame(table(galton$child, galton$parent))
names(freqData) <- c("child","parent","freq")
freqData$child <- as.numeric(as.character(freqData$child))
freqData$parent <- as.numeric(as.character(freqData$parent))
g <- ggplot(filter(freqData, freq > 0), aes(x = parent, y = child)) + 
  scale_size(range = c(2,10), guide = "none") + 
  geom_point(colour = "grey50", aes(size = freq+10)) + 
  geom_point(aes(colour = freq, size = freq)) + 
  scale_colour_gradient(low = "lightblue", high = "white")
g
```

Here's cleaner code to get the frequency data. This is also an alternative to plotting the information:

```{r}
freqData2 <- galton %>% 
  group_by(child, parent) %>% 
  summarise(freq = n())
g2 <- ggplot(galton, aes(x = parent, y = child)) +
  geom_jitter()
g3 <- ggplot(freqData2, aes(x = parent, y = child)) +
  scale_size(range = c(2,10), guide = "none") + 
  geom_point(colour = "grey50", aes(size = freq+10)) + 
  geom_point(aes(colour = freq, size = freq)) + 
  scale_colour_gradient(low = "lightblue", high = "white")
g2
g3
```

## Verify hand calculations and R calculations are the same for the coefficients
```{r}
y <- galton$child
x <- galton$parent
beta1 <- cor(y,x) * sd(y)/sd(x)
beta0 <- mean(y) - beta1* mean(x)
coef(lm(y ~ x))
c(beta0, beta1)
```

## Regression through the origin yields the same slope as linear regression if you mean center the y's and x's
```{r}
yc <- y - mean(y)
xc <- x - mean(x)
beta1 <- sum(xc * yc)/sum(xc^2)
c(beta1, coef(lm(y ~ x))[2])
#get linear model without intercept
lm(yc ~ xc - 1)
```

## Normalizing the data results in the slope being the correlation
```{r}
yn <- (y - mean(y))/sd(y)
xn <- (x - mean(x))/sd(x)
c(sd(yn),sd(xn)) #check to make sure sd is 1
c(cor(y,x), cor(yn,xn), coef(lm(yn ~ xn))[2])
```

## Add the regression line with the confidence level
```{r}
g + geom_smooth(method = "lm", formula = y ~ x)
```