---
title: "Analysis of HollywoodMovies Dataset"
author: "Santosh Jnawali, Rizon Giri, Kevil Khadka"
date: "11/7/2019"
output:
  slidy_presentation: default
  ioslides_presentation: default
  beamer_presentation: default
Course: STAT-361
Assignment: Midterm Project
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(GGally)
library(tidyverse)
library(dplyr)
library(knitr)
```

## Introduction
- Dataset : HollywoodMovies2011

- Source : Lock5Data package

- Question of interest : Are there any possible variables that could help us to predict the success of the movies?

- Method applied : Linear Regression 

## Data

Our dataset had 14 variables for 136 movies and since we only wanted to know the most significant predictor variables which could significantly predict the world gross of the movies, we had to eliminate some of the variables which was of least importance to our model and choose only significant predictors. In order to have a general idea of which variables to select in the model, we first ran a scatterplot matrix for all the variables. 

## Matrix of scatterplot and taking out the important variables for prediction
```{r echo= FALSE}
hollywood<- Lock5Data::HollywoodMovies2011

project<- hollywood %>%
        dplyr::select(Genre,Budget,OpeningWeekend,WorldGross)

hollywood1<-project%>%
     filter(Genre %in% c("Comedy","Action","Drama"))
  newdata<-na.omit(hollywood1) %>%
    arrange()
```

```{r, out.width='75%', fig.align='center'}
    ggpairs(data=newdata,columns = c(4,3,2))
```
- Looking at the scatterplot matrix, it seems OpeningWeekend and Budget can explain the Worldgross after fitting a linear model.

## Scatterplot of WorldGross and Budget
```{r,out.width='80%', fig.align='center'}
ggplot(data=newdata,aes(Budget,WorldGross))+
  geom_point(aes(color=Genre))+
  geom_smooth(method = "lm")
```
- The scatterplot of WorldGross and Budget shows a pretty linear relationship between the variables.

## Scatterplot of WorldGross and Opening Weekend 
```{r,out.width='80%', fig.align='center'}
ggplot(data=newdata,aes(OpeningWeekend,WorldGross))+
  geom_point(aes(color=Genre))+
  geom_smooth(method = "lm")
```
- The scatterplot of OpeningWeekend and Worldgross also seems to be linear.

## Plot of WorldGross,OpeningWeekened, and Budget
```{r,out.width='80%', fig.align='center'}
ggplot(data =newdata, mapping = aes(x = OpeningWeekend , y = WorldGross, size=Budget)) +
  geom_point(mapping = aes(color = Genre))
```
- From the combined plot, we can see that with increase in openingWeekend, there seems to be increase in Budget with WorldGross. 
So, the combination of the plot, gives us idea that, there is linear relationship between these variable.

## Preliminary Analysis
- We started our analysis from fitting very simple regression model to get an idea of what's happening in the dataset

## Fitting the multiple regression of WorldGross on Budget and OpeningWeekend
- After fitting several simple linear models with individual variable OpeningWeekend and Budget, checking assumptions and dealing with outliers, we thought we should go with multiple regression model.
- Thus,we decided to fit the multiple regression. To begin, we regressed WorldGross on Budget and OpeningWeekend.

```{r, out.width='70%', fig.align='center'}
fit4<- lm((WorldGross)~(Budget)+(OpeningWeekend),data=newdata)
summary(fit4)
```

## Diagnostic Plots 
- Looking at the residual plot, there seems to be some fanning. The qqplot seems pretty good and the histogram is fairly normal. In order to deal with our assumption, we fit our model with log transformation. 
```{r, out.width='70%', fig.align='center'}
plot(fit4,1)
plot(fit4,2)
hist(fit4$residuals)
```

## Transformation model
- After trying several log transformation of variables, we ended of transforming variable, WorldGross and OpeningWeekend as it fits our assumptions.
- Summary of the model "Fit5"
```{r, out.width='70%', fig.align='center'}
fit5<- lm(log(WorldGross)~(Budget)+log(OpeningWeekend),data=newdata)
summary(fit5)
```

## Plotting the model after Log transformation
- Looking at these diagnostic plots after transforming the data, we can see the residual plot seems pretty good. 
- the data points are held in the middle of the qqplot and histogram of the residuals seems pretty normal too.
- Therefore we can assume our regressions model fits the assumptions and can be used for further analysis.

```{r, out.width='70%', fig.align='center'}
ggplot(data = fit5) +
  geom_point(mapping = aes(x = .fitted, y = .resid)) +
  geom_hline(yintercept = mean(fit5$residuals))
```

```{r, out.width='70%', fig.align='center'}
plot(fit5,2)
hist(fit5$residuals)
```

## Summary of the Model with Genre:
- However, we have not included our categorical variable Genre in that model, so we again want to fit a model using all the available variable which includes additional variable 'Genre'.
```{r, out.width='70%', fig.align='center'}
fit6<- lm(log(WorldGross)~(Budget)+log(OpeningWeekend)+factor(Genre),data=newdata)
summary(fit6)
```

## Plotting residual vs fitted value on log transformation of variables with factor Genre
- Looking at these diagnostic plots after transforming the data, we can see the residual plot seems pretty good.
- QQplot and histogram of the residuals seem pretty normal. 
- we can conclude that this regression model also fits the assumptions, and use for further analysis.

```{r, out.width='70%', fig.align='center'}
plot(fit6,1)
plot(fit6,2)
hist(fit6$residuals)
```


## Analysis and Refining Model
- Based on our question of interest, we want to predict the WorldGross of a movie based on the different available variables.
- We have fitted several regression using those variable.
- We want to make our regression as simpler as we can, so we are now comparing our reduced model "fit5" and full model "fit6" to know with which model should we proceed. 
- * "fit5" - multiple regression model of log_WorldGross on log_OpeningWeekend and Budget
- * "fit6" - multiple regression model of log_WorldGross on log_OpeningWeekend and Budget including Genre as Factor

## Comparison of Full and reduced model 
- Our hypotheses are stated below:
- Null hypothesis, $H_0$ : There is no lack of fit
- Alternative Hypothesis, $H_a$ : There is lack of fit

- ANOVA table 
```{r, out.width='50%', fig.align='center'}
anova(fit5,fit6)
```
- The comparison of anova shows the p-value greater than 0.05.
- There is not sufficient evidence to reject our null hypothesis.This means there is no lack of fit. 
- So we can proceed with our simple regression of model fit5 for our analysis. 
- Thus, our model with Budget and log(Openingweekend) can explain as much as the model created with additional variable Genre. There is no significant effect of Genre found in the linear model, so we can use model fit5 for our analysis and use it for the prediction. 

## Final Regression Model
```{r, out.width='70%', fig.align='center'}
summary(fit5)
```
- Our Final regression model is given by 

 \mu\{{log(WorldGross)}|{Budget},{log(OpeningWeekend)}} 
= 2.283746 +0.007003 * Budget + 0.658244 * log(OpeningWeekend)


## Conclusion
  After conducting the complete analysis of the data, we reached to our conclusion that, our model "fit5" in which response variable WorldGross and predictor variable OpeningWeekend are logged, is better in predicitng the financial success of the movies, based upon the variables Budget and OpeningWeekend. In our final model "fit5", predictor variables, log of OpeningWeekend and Budget explained about 81.24 % variability in reponse variable, log of WorldGross. Over the time of the analysis, we realized that there were other variables in the dataset which were excluded during the analysis. There might have been some kind of interaction effect between those variables which we were not able to conduct within the given time frame of project. We would have came up with better model if we were able to do so. However, in the conducted analysis, we found model "fit5" better in predicting our response.